Compositional Certified Resource Bounds

### 2014-07-03 paper/main.tex
experiments

To compare our tool with existing work, we focused on loop bounds and
use a simple metric that counts the number of back edges (i.e., number
of loop iterations) that are follows in the execution of the program
because most other tools only bound this specific cost. In
\pref{fig:compar}, we show the bounds we derived (AAA) together with
the bounds derived by LOOPUS~\cite{SinnZV14} and
Rank~\cite{AliasDFG10}.  We also contacted the authors of SPEED but
have not been able to obtain this tool.
KoAT~\cite{BrockschmidtEFFG14} and PUBS~\cite{AlbertAGPZ12} currently
cannot operate on C code and the examples would need to be manually
translated into a term-rewriting to be analyzed by these tools.  For
Rank it is not totally clear how the computed bound relates to the C
program since the computed bound is for transitions in an automaton
that is derived from the C code.  For instance, the bound $2+y-x$ that
is derived for \emph{t08}.

@@

\begin{figure}
\centering
\begin{tabular}{r|c|c|c|c|c}
& KoAT & Rank & LOOPUS & SPEED & AAA \\
\hline
\#bounds & 9 & 24 & 20 & 14 & 32
\\
\#lin.\ bounds & 9 & 21 & 20 & 14 & 32
\\
\#best bounds & 0 & 0 & 11 & 14 & 29
\\
\#tested & 14 & 33 & 33 & 14 & 33
\end{tabular}
\caption{Comparison of our automatic amortized analysis with other
  automatic tools.  We have not been able to obtain a version of
  SPEED~\cite{GulwaniMC09} and just use the bounds that have been
  reported by the authors. Similarly, KoAT~\cite{BrockschmidtEFFG14}
  does currently not work on C programs and we use the bounds that
  have been reported in the author's experimental evaluation.}
\label{fig:compar}
\end{figure}

@@

\begin{figure*}
\small

@@

%--------------------------------------------
\end{tabular}
\caption{
  Experimental evaluation comparing the bounds generated KoAT, Rank,
  Loopus, Speed, and our automatic amortized analysis on several
  challenging linear examples. Results for KoAT and Speed were extracted
  from older publications because KoAT cannot take C programs as 
  input in its current version and Speed is not publically available.
  Entries marked with ? signal that we cannot test the tool. Entries
  marked with --- mean that the tool failed to produce a result.
}
\label{fig:eval}
\end{figure*}

>>>>

To compare our tool with existing work, we focused on loop bounds and
use a simple metric that counts the number of back edges (i.e., number
of loop iterations) that are follows in the execution of the program
because most other tools only bound this specific cost. In
\pref{fig:compar}, we show the bounds we derived (AAA) together with
the bounds derived by LOOPUS~\cite{SinnZV14} and
Rank~\cite{AliasDFG10}.  We also contacted the authors of SPEED but
have not been able to obtain this tool.
KoAT~\cite{BrockschmidtEFFG14} and PUBS~\cite{AlbertAGPZ12} currently
cannot operate on C code and the examples would need to be manually
translated into a term-rewriting to be analyzed by these tools.  For
Rank it is not totally clear how the computed bound relates to the C
program since the computed bound is for transitions in an automaton
that is derived from the C code.  For instance, the bound $2+y-x$ that
is derived for \emph{t08} only applies to the first loop in the
program.

\pref{tab:compar} summarizes the results of our experiments.  It shows
for each tool the number of derived bounds (\#bounds), the number of
asymptotically tight bounds (\#lin. bounds), the number of bounds with
the best constant factors in comparison with the other tools (\#best
bounds), and the number of examples that we were able to test with the
tool (\#tested).  Since we were not able to run the exponents for KoAT
and SPEED, we simply used the bounds the have been reported by the
authors of the respective tools.  The results show that our automatic
amortized analysis outperforms the existing tools on our example
programs.  However, this experimental evaluation has to be taken with
a grain of salt.  Our goal in this work is not to set a new standard
for automatic bound analysis but only to show that our approach has
advantages on certain types of examples.  Overall the existing tools
are more powerful since they can derive polynomial bounds and support
more features of C.  We were particularly impressed by LOOPUS which is
very robust, works on large C files, and derives very precise bounds.
We did not include the running times of the tools in the table since
all the tested tools work very efficiently and need less then a second
on all tested examples.

Since we have a formal cost semantics, we can run our examples with
this semantics for different inputs and measure the cost to compare it
to our derived bound.  \pref{fig:3d} shows such a comparison for
example \emph{t08}.  One can see the derived constant factors are the
best possible if the input variable $x$ is non-negative.  If $x$ is
negative then the bound is only slightly off.

@@

\begin{table}
\centering
\begin{tabular}{r|c|c|c|c|c}
& KoAT & Rank & LOOPUS & SPEED & AAA \\
\hline
\#bounds & 9 & 24 & 20 & 14 & 32
\\
\#lin.\ bounds & 9 & 21 & 20 & 14 & 32
\\
\#best bounds & 0 & 0 & 11 & 14 & 29
\\
\#tested & 14 & 33 & 33 & 14 & 33
\end{tabular}
\caption{Comparison of our automatic amortized analysis with other
  automatic tools.  We have not been able to obtain a version of
  SPEED~\cite{GulwaniMC09} and just use the bounds that have been
  reported by the authors. Similarly, KoAT~\cite{BrockschmidtEFFG14}
  does currently not work on C programs and we use the bounds that
  have been reported in the author's experimental evaluation.}
\label{tab:compar}
\end{table}

@@

\begin{table*}
\small

@@

%--------------------------------------------
\end{tabular}
\caption{
  Experimental evaluation comparing the bounds generated KoAT, Rank,
  Loopus, Speed, and our automatic amortized analysis on several
  challenging linear examples. Results for KoAT and Speed were extracted
  from older publications because KoAT cannot take C programs as 
  input in its current version and Speed is not publically available.
  Entries marked with ? signal that we cannot test the tool. Entries
  marked with --- mean that the tool failed to produce a result.
}
\label{fig:eval}
\end{table*}
